Temperature, wave exposure and wind as drivers of morphology in two kelp species

### 2019-07-14 paper_morph_new.Rmd
script for deep ecklonia and laminaria added. Final outputs need to be updated. RDA to follow

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
```

## Introduction

@@

## Methods
### Site selection

@@

## Results
### Temperature variables

@@

## Preparing data
# Complete lon/lats
kelp_sites <- read_csv("metadata/sites_complete.csv")

# Ecklonia
eck_data <- read_csv("data/Ecklonia_morphometrics.csv") %>%
  mutate(site = replace(site, site == "Buffels Bay", "Buffelsbaai")) %>%
  mutate(site = replace(site, site == "St. Helena Bay", "St Helena Bay")) %>%
  # Not using Rocky Bank in this study
  filter(site != "Rocky Bank")

# Laminaria
lam_data <- read_csv("data/Laminaria_morphometrics.csv") %>%
  mutate(site = replace(site, site == "Buffels Bay", "Buffelsbaai")) %>%
  # NB: Even though I am changing the name, there will still be no lon/lat values for this site
  # We need to decide how we want to handle this. Give it the same lon/lat as Buffelsbaai?
  mutate(site = replace(site, site == "Buffels Bay South", "Buffelsbaai South")) %>%
  # NB: This may also not be an appropriate change
  mutate(site = replace(site, site == "Bordjies reef North", "Bordjies Deep")) %>%
  # Not using Rocky Bank in this study
  filter(site != "Rocky Bank")

@@

# The SACTN monthly data
load("data/SACTNmonthly_v4.1.Rdata")

# The missing site temperatures
load("data/interp_sites.Rdata")

@@

```{r Temperature variable processing, message=FALSE, warning=TRUE, include=FALSE}

@@

```{r Temperature variables plot, echo=FALSE, include = FALSE, fig.cap="Temperature variables at the collection sites around the Cape Peninsula. Temperature variables include minimum represented by blue dots.", fig.height=9, fig.width=7}

@@

### Wave variables

```{r Wave data prep, message=FALSE, warning=FALSE, include=FALSE}

@@

# The site list key
wave_sites <- read_csv("metadata/sites_complete.csv") %>% 
  select(site, wave_7, wave_15)

@@

```{r Wave variable processing, warning=FALSE, include=FALSE, echo = FALSE}

@@

```{r Wave variable plot, echo=FALSE, fig.cap='Wave variables at each collection site around the Western Cape coast. Mean variables are represented by dots and standard deviation by whiskers. Plot "A" represents annual significant wave height.', fig.height=8, fig.width=7}

@@

>>>>

# Introduction

@@

# Methods
### Site selection

@@

# Results
## Temperature variables

@@

## Preparing data
# Complete lon/lats
kelp_sites <- read.csv("metadata/site_meta_complete.csv", sep = ";")

# Ecklonia
eck_data <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
  filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
  filter(site!= "De Hoop")

# Laminaria
lam_data <- read.csv("data/morph_data/lam_morph_data.csv", sep = ";") %>%
  mutate(site = replace(site, site == "Buffels Bay", "Buffelsbaai")) %>%
  # NB: Even though I am changing the name, there will still be no lon/lat values for this site
  # We need to decide how we want to handle this. Give it the same lon/lat as Buffelsbaai?
  mutate(site = replace(site, site == "Buffels Bay South", "Buffelsbaai South")) %>%
  # NB: This may also not be an appropriate change
  mutate(site = replace(site, site == "Bordjies reef North", "Bordjies Deep")) %>%
  # Not using Rocky Bank in this study
  filter(site != "Rocky Bank")

@@

# The SACTN monthly data
load("data/temp_data/SACTNmonthly_v4.1.Rdata")

# The missing site temperatures
load("data/temp_data/interp_sites.Rdata")

@@

```{r Temperature variable processing, message=FALSE, warning=FALSE, include=FALSE}

@@

```{r Temperature variables plot, echo=FALSE, warning=FALSE, fig.cap="Temperature variables at the collection sites around the Cape Peninsula. Temperature variables include minimum represented by blue dots.", fig.height=9, fig.width=7}

@@

## Wave variables

```{r wave data prep, message=FALSE, warning=FALSE, include=FALSE}

@@

# The site list key
wave_sites <- read_csv2("metadata/site_meta_complete.csv") %>% 
  select(site, wave_7, wave_15)

@@

```{r wave variable processing, warning=FALSE, include=FALSE, echo = FALSE}

@@

```{r wave variable plot, warming = FALSE, echo=FALSE, message = FALSE, include= TRUE, fig.cap='Wave variables at each collection site around the Western Cape coast. Mean variables are represented by dots and standard deviation by whiskers. Plot "A" represents annual significant wave height.', fig.height=8, fig.width=7}

@@

## Morphology
### *L. pallida*
```{r laminaria processing, message=FALSE, warning=FALSE, include=FALSE}

lam_data <- read.csv("data/morph_data/lam_morph_data.csv", sep = ";") %>% 
   filter(site!= "Rocky Bank")

# remove unnecessary columns
lam_df <- subset(lam_data, select = -c(1,2,6))

pos_lam <- c("Betty's Bay", "Roman Rock","A-Frame", "Miller's Point", "Baboon Rock", 
                  "Batsata Rock", "Buffels Bay"," Buffels Bay South","Bordjies North","Olifantsbos", "Kommetjie", "Sea Point")
pos_exposure_lam <- c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Sheltered", "Moderatley exposed", "Exposed", "Very exposed")

lamina_weight_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaWeight, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
lamina_length_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaLength, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
lamina_thickness_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = laminaThickness, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
stipe_mass_lam_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeMass, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
stipe_length_lam_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeLength, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
stipe_diameter_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = stipeDiameter, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
digits_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = digits, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
thallus_mass_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = thallusMass, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
total_length_plot <- ggplot(data = lam_df, aes(x = wave_exposure, y = totalLength, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure_lam) + 
         geom_boxplot(outlier.shape = NA) 
```

```{r laminaria plots, echo=FALSE, message=FALSE, warning=FALSE}
grid.arrange(lamina_weight_plot,lamina_length_plot,lamina_thickness_plot,stipe_mass_lam_plot,stipe_length_lam_plot,stipe_diameter_plot, digits_plot, thallus_mass_plot,total_length_plot,total_length_plot, ncol = 3, nrow = 4)

### *E.maxima*

```{r ecklonia  processing, message=FALSE, warning=FALSE, include=FALSE}

library(ggplot2)

## Preparing data
# Ecklonia
eck_data <- read.csv("data/morph_data/eck_morph_data.csv", sep = ";") %>%
 filter(site!= "Rocky Bank") %>%   # Not using Rocky Bank in this study
 filter(site!= "De Hoop")  
# group_by(wave_exposure, frond_mass) %>% 
# summarise(mean = mean(frond_mass:total_length, na.rm = TRUE)) %>% 
# ungroup()

eck_deep <- eck_data %>% 
  filter(population == "Deep")

# custom positions for wave exposure categories and sites
pos_exposure <- c("Fully sheltered", "Extremely sheltered", "Very sheltered", "Sheltered", "Moderatley exposed", "Exposed", "Very exposed")
pos_eck<- c("Betty's Bay", "Kalk Bay", "A-Frame",  "Miller's Point", "Baboon Rock", "Batsata Rock", "Buffelsbaai", "Scarborough", "Soetwater","Kommetjie","Hout Bay", "Oudekraal", "St Helena Bay")

frond_mass_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = frond_mass, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure) + 
         geom_boxplot(outlier.shape = NA) 
primary_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_length, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure) + 
         geom_boxplot(outlier.shape = NA) 
primary_width_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = primary_width, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure) + 
         geom_boxplot(outlier.shape = NA) 
frond_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = frond_length, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure) + 
         geom_boxplot(outlier.shape = NA) 
stipe_mass_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_mass, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure) + 
         geom_boxplot(outlier.shape = NA) 
stipe_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_length, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure) + 
         geom_boxplot(outlier.shape = NA) 
stipe_circ_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = stipe_circ, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure) + 
         geom_boxplot(outlier.shape = NA) 
tufts_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = tufts, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure) + 
         geom_boxplot(outlier.shape = NA) 
epi_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = epi_length, fill = wave_exposure)) +
         scale_x_discrete(limits = pos_exposure) + 
         geom_boxplot(outlier.shape = NA) 
total_length_plot <- ggplot(data = eck_deep, aes(x = wave_exposure, y = total_length, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure) + 
  geom_boxplot(outlier.shape = NA) +
  ylim(0, 2000)

```{r ecklonia plots, echo=FALSE, message=FALSE, warning=FALSE}

grid.arrange(frond_mass_plot,primary_length_plot,primary_width_plot,frond_length_plot,stipe_mass_plot,stipe_length_plot,stipe_circ_plot, tufts_plot, epi_length_plot,total_length_plot, ncol = 3, nrow = 4)
